---
title: data.table reverse dependency checks
output: html_document
---

```{r Ropts, echo=FALSE}
options(width=200)
knitr::opts_chunk$set(
  echo=FALSE,
  results="asis")
```

```{r results=TRUE}
plural <- function(number, singular){
  s <- ifelse(number==1, "", "s")
  sprintf("%d %s%s", number, singular, s)
}
suppressMessages({
  if(!requireNamespace("nc"))install.packages("nc")
  if(!requireNamespace("slurm")){
    install.packages("~/R/slurm", repos=NULL)
  }
  library(data.table)
})
job.dir <- readLines("JOBDIR",n=1)
job.date <- basename(job.dir)
deps.dt <- data.table::fread(file.path(job.dir, "deps.csv"))
JOBID.file <- file.path(job.dir, "JOBID")
JOBID <- readLines(JOBID.file, n=1)
sacct.arg <- paste0("-j", JOBID)
sacct.dt <- slurm::sacct(sacct.arg)
some.sacct <- sacct.dt[, .(
  task, time=Elapsed, MB=round(megabytes), State=State_blank
)]
deps.sacct <- some.sacct[deps.dt, on=.(task=task.id)]
dt.tar.gz.vec <- Sys.glob(file.path(job.dir, "*.tar.gz"))
dt.version.vec <- gsub(".tar.gz$|.*data.table_", "", dt.tar.gz.vec)

## Read all logs.
log.file.glob <- file.path(
  job.dir, "packages", "*", "log.txt")
grep.dep.cmd <- paste(
  "grep '^also installing the dependencies'",
  log.file.glob)
grep.dep.out <- system(grep.dep.cmd, intern=TRUE)
log.pattern <- list(
  log.txt=list(
    ".*/",
    checked=".*?",
    "/log.txt"),
  ":")
grep.dep.dt <- nc::capture_first_vec(
  grep.dep.out,
  log.pattern,
  "also installing the dependencies ",
  deps.str=".*")
also.dt <- nc::capture_all_str(
  grep.dep.dt$deps.str,
  "'",
  dep=".*?",
  "'")
also.counts <- also.dt[, .(downloads=.N), by=dep][order(-downloads)]
if(FALSE){
  data.table::fwrite(also.counts[1:200], "popular_deps.csv")
}

## Read all diffs.
diff.file.vec <- Sys.glob(file.path(
  job.dir, "packages", "*", "*", "significant_differences.csv"))
diff.file.cols <- names(fread(diff.file.vec[1], nrows=0))
sig.diff.wide <- data.table(diff.csv=diff.file.vec)[, {
  fread(diff.csv, colClasses=list(character=diff.file.cols))
}, by=diff.csv]
Rversion.pattern <- list(
  job.dir,
  "/packages/",
  Package=".*?",
  "/",
  Rversion=".*?",
  "/")
sig.cols <- c(
  "first.bad.commit",
  "Package", "checking", "Rvers",
  "comments",
  "CRAN", "release", "master")
sig.diff.dt <- nc::capture_first_df(
  sig.diff.wide, diff.csv=Rversion.pattern
)[, Rvers := ifelse(grepl("devel", Rversion), "devel", "release")
  ][, sig.cols, with=FALSE]
setkeyv(sig.diff.dt, sig.cols)
deps.sacct[, sig.diffs := NA_integer_]
deps.sacct[sig.diff.dt, sig.diffs := .N, by=Package, on=.(Package)]
deps.sacct[is.na(sig.diffs), sig.diffs := 0]

## add number of missing check logs.
check.file.glob <- file.path(
  job.dir, "packages", "*", "*", "*", "00check.log")
check.file.vec <- Sys.glob(check.file.glob)
check.file.meta <- nc::capture_first_vec(
  check.file.vec,
  check.log=list(
    Rversion.pattern,
    dt.version=".*?",
    "/00check.log"))
CJ.args <- lapply(check.file.meta[, .(Rversion, dt.version)], unique)
CJ.args$Package <- deps.dt$Package
select.dt <- do.call(CJ, CJ.args)
expected.checks <- check.file.meta[select.dt, on=names(select.dt)]
expected.checks[, dt.vers := sub("_.*", "", dt.version)]
expected.checks[, log.txt := file.path(job.dir, "packages", Package, "log.txt")]
missing.logs <- expected.checks[is.na(check.log)]
deps.sacct[, no.log := NA_integer_]
deps.sacct[missing.logs, no.log := .N, by=Package, on=.(Package)]
deps.sacct[is.na(no.log), no.log := 0]

## report versions used.
pvers <- function(what, ver.vec){
  cat(what, "versions checked:\n")
  print(ver.vec)
}
pvers("R", unique(check.file.meta$Rversion))
pvers("data.table", dt.version.vec)

## add counts of config and dl failed.
grep.config.fail <- paste(
  "grep 'configuration failed for package'",
  log.file.glob)
config.fail.vec <- suppressWarnings({
  system(grep.config.fail, intern=TRUE)
})
deps.sacct[, config.fail := NA_integer_]
if(length(config.fail.vec)){
  config.fail.uniq <- unique(config.fail.vec)
  config.fail.dt <- nc::capture_first_vec(
    config.fail.uniq,
    log.pattern,
    "ERROR: configuration failed for package '",
    failed=".*?",
    "'")
  config.fail.first <- config.fail.dt[, .SD[1], by=failed]
  deps.sacct[
    config.fail.dt,
    config.fail := .N,
    by=Package,
    on=.(Package=checked)]
}else{
  config.fail.first <- data.table(failed=character(), checked=character())
}
deps.sacct[is.na(config.fail), config.fail := 0]
grep.dl.fail <- paste(
  "grep 'download of package.*failed'",
  log.file.glob)
dl.fail.vec <- suppressWarnings({
  system(grep.dl.fail, intern=TRUE)
})
deps.sacct[, dl.fail := NA_integer_]
if(length(dl.fail.vec)){
  dl.fail.dt <- nc::capture_first_vec(
    dl.fail.vec,
    log.pattern,
    ".*?download of package '",
    failed=".*?",
    "' failed")
  deps.sacct[dl.fail.dt, dl.fail := .N, by=Package, on=.(Package=checked)]
}
deps.sacct[is.na(dl.fail), dl.fail := 0]

## Add column to show if diff is new since last run.
sig_diff_files <- grep(
  job.date,
  Sys.glob("analyze/*/significant_differences.csv"),
  invert=TRUE, value=TRUE)
prev.sig_diff.csv <- sig_diff_files[length(sig_diff_files)]
prev.sig_diff.dt <- fread(prev.sig_diff.csv)
diff.join.cols <- c(
  "Package", "checking", "Rvers", "release", "master")
newly.fixed <- prev.sig_diff.dt[!sig.diff.dt, on=diff.join.cols]
sig.diff.dt[, new := "NEW!"]
sig.diff.dt[prev.sig_diff.dt, new := "", on=diff.join.cols]

## copy log files from scratch to projects/analyze subdir.
log.file.vec <- Sys.glob(log.file.glob)
log.pkg <- basename(dirname(log.file.vec))
analyze.dir <- file.path("analyze", job.date)
dir.create(analyze.dir, showWarnings=FALSE)
pkg.txt.vec <- file.path(analyze.dir, paste0(log.pkg, ".txt"))
copied <- file.copy(log.file.vec, pkg.txt.vec)
ahref <- function(href, text){
  sprintf('<a href="%s">%s</a>', href, text)
}
mytab <- function(DT){
  copied <- data.table(DT)
  if("CRAN" %in% names(copied)){
    copied[, CRAN := ahref(sprintf(
      "https://cloud.r-project.org/web/checks/check_results_%s.html",
      Package),
      CRAN)]
  }
  if("first.bad.commit" %in% names(copied)){
    copied[, first.bad.commit := paste(ahref(paste0(
      "https://github.com/Rdatatable/data.table/commit/",
      first.bad.commit),
      first.bad.commit),
      comments)]
    copied[, comments := NULL]
  }
  show.cols <- c("sig.diffs","no.log","config.fail","dl.fail")
  some.cols <- show.cols[show.cols %in% names(copied)]
  if(length(some.cols)){
    int.mat <- as.matrix(copied[, some.cols, with=FALSE])
    name.mat <- matrix(some.cols, nrow(int.mat), ncol(int.mat), byrow=TRUE)
    chr.mat <- ifelse(int.mat==0, "", sprintf("%s=%d", name.mat, int.mat))
    copied[, warnings := apply(chr.mat, 1, paste, collapse=" ")]
    set(copied, j=some.cols, value=NULL)
  }
  copied[, Package := ahref(paste0(
    Package, ".txt"),
    Package)]
  knitr::kable(copied)
}

count.dt <- deps.sacct[, .(
  dep.pkgs=.N
), keyby=.(
  COMPLETED=State=="COMPLETED",
  diffs=sig.diffs>0,
  fail=config.fail|dl.fail
)]
digits.pattern <- list("[0-9]{2}", as.integer)
time.dt <- nc::capture_first_vec(
  deps.sacct$time,
  hours.only=digits.pattern, ":",
  minutes.only=digits.pattern, ":",
  seconds.only=digits.pattern
)[,
  minutes := hours.only*60 + minutes.only + seconds.only/60
  ]
time.wide <- dcast(
  time.dt,
  . ~ .,
  list(min, median, max, sum),
  value.var="minutes"
)
time.tall <- nc::capture_melt_single(
  time.wide[,-1],
  "minutes_",
  stat=".*",
  value.name="minutes"
)[, hours := minutes/60][, days := hours/24][]
source("myStatus.R")#for get_flavor
```

`r plural(nrow(deps.sacct), "package")` checked, time stats:

```{r}
knitr::kable(time.tall, digits=2)
```

## `r plural(nrow(sig.diff.dt), "Significant difference")` found

The table below has one row for each check found to have different
results when using two versions of `data.table` (master and
release). The current CRAN result for that check is also shown (using
linux platform with either `r get_flavor("r-devel")` or `r get_flavor("r-release")`).
Instructions about how to interpret and file issues based on these results are
explained [on a wiki page](https://github.com/Rdatatable/data.table/wiki/Release-management-and-revdep-checks).


```{r}
mytab(sig.diff.dt)
```

## `r plural(nrow(newly.fixed), "significant difference")` fixed since last check

The significant differences below were found in the last check, `r prev.sig_diff.csv` but are no longer present in the current check above. 
If they are really fixed, then please search for the corresponding [issue](https://github.com/Rdatatable/data.table/issues) and close it.

```{r}
mytab(newly.fixed)
```

## `r plural(nrow(config.fail.first), "Configuration failure")`

In the table below, the `failed` column shows which package failed to
configure/install, and the `Package` column shows the log file of the
package that was supposed to be checked. Each failure is only reported
once (only the first log with that configuration failure is
shown). Ideally this table should be empty, so if there are entries
here, they should be reported to the maintainer of this revdep check
system via [this
issue](https://github.com/Rdatatable/data.table/issues/5546).

```{r}
mytab(config.fail.first[, .(failed, Package=checked)])
```

## Longest running jobs

These are the top 10 packages in terms of time to install, then
compute checks and git bisect. If any have `State=TIMEOUT` regularly,
then there are two options to fix:

* check the log, and if there is a lot of time spent installing packages, then this package should be added to [popular_deps.csv](https://github.com/tdhock/data.table-revdeps/blob/master/popular_deps.csv), which will increase time spent installing packages to the shared library prior to launching check jobs, but will decrease the time of each check job that would have had to install the popular packages to the job-specific library.
* otherwise, if not much time is spent installing packages, then increase the `--time` parameter of the `run_one.sh` script in [params.R](https://github.com/tdhock/data.table-revdeps/blob/master/params.R).

```{r}
mytab(deps.sacct[order(-time)][1:10])
```

## Largest memory jobs

These are the top 10 packages in terms of max memory usage in the job
which installs, then computes checks and git bisect. If any go over
the max and are killed, then the `--mem` parameter of the `run_one.sh`
script should be increased in
[params.R](https://github.com/tdhock/data.table-revdeps/blob/master/params.R).

```{r}
mytab(deps.sacct[order(-MB)][1:10])
```

## Most installed packages

These are the top 10 packages which appeared in "also installing the
dependencies..." in the logs. Packages in this table should be
considered for inclusion in
[popular_deps.csv](https://github.com/tdhock/data.table-revdeps/blob/master/popular_deps.csv),
which is the list of packages that are installed to the shared
library, prior to launching the check jobs (each of which has its own
job-specific library where packages are installed if they are not
present in the shared library).

```{r}
knitr::kable(also.counts[1:min(10,.N)])
```

## Full list of jobs

The SLURM job ID was `r JOBID`, and the task ID is given in the first
column of the table below, which has a row for every reverse
dependency. If any of the following are not zero, then there will be
an entry in the warnings column:

* `sig.diffs`: rows in the significant differences table.
* `no.log`: missing log files (there could be up to 4).
* `config.fail`: dependent packages which failed to configure/build.
* `dl.fail`: packages which failed to download.

```{r}
mytab(deps.sacct)
```

## CSV data

```{r}
fsave <- function(DT, f){
  out.csv <- paste0(f, ".csv")
  out.path <- file.path(analyze.dir, out.csv)
  data.table::fwrite(DT, out.path)
  cat('<p>', ahref(out.csv, f), '</p>\n')
}
fsave(deps.sacct, "full_list_of_jobs")
fsave(sig.diff.dt, "significant_differences")
```

## Source code

Computed using [code](https://github.com/tdhock/data.table-revdeps)
written by Toby Dylan Hocking.
